Medical Imaging Classification with Cold-Atom Reservoir Computing using Auto-Encoders and Surrogate-Driven Training\\

### 2025-05-27 paper/main.tex
advancements

\begin{equation}
    \mathcal{L} = (1-\lambda) \cdot \mathcal{L}_R + \lambda \cdot \mathcal{L}_{C},
\end{equation}
where:
\begin{itemize}
    \item \( \mathcal{L}_R = \| x - \hat{x} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( x \) and its reconstruction \( \hat{x} \).
    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \).
    \item \( 0 < \lambda < 1 \) is a hyperparameter that controls the trade-off between reconstruction and classification objectives.
\end{itemize}

@@

%============================================
% EXPERIMENTAL SETUP
%============================================
\section{Experimental Setup}
\subsection{Datasets}
\subsection{Implementation Details}
\subsubsection{Quantum Simulation Parameters}
\subsubsection{Classical Network Architectures}
\subsection{Comparison Methods}
\subsection{Performance Metrics}
\subsection{Parameter Sweep Strategy}

>>>>

\begin{equation}
    \mathcal{L} = (1-\lambda) \cdot \mathcal{L}_R + \lambda \cdot \mathcal{L}_{C},
    \label{eq:composed_loss_function}
\end{equation}
where:
\begin{itemize}
    \item \( \mathcal{L}_R = \| x - \hat{x} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( x \) and its reconstruction \( \hat{x} \).
    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \).
    \item \( 0 < \lambda < 1 \) is a hyperparameter that controls the trade-off between reconstruction and classification objectives.
\end{itemize}

@@

%============================================
% EXPERIMENTAL SETUP
%============================================
\section{Experimental Setup}
\subsection{Datasets}
We evaluate our proposed architecture on three primary datasets:
\begin{enumerate}
    \item \textbf{Synthetic Polyp Dataset:} 
    A synthetic dataset of polyp images generated to simulate 
    realistic medical imaging scenarios.

    \item \textbf{CVC-ClinicDB Dataset:} 
    Real image patches extracted from the CVC-ClinicDB 
    dataset, a well-known benchmark for polyp detection. 
    This dataset provides a diverse set of polyp images 
    with varying characteristics, allowing for robust 
    evaluation of classification performance.

    \item \textbf{MNIST Dataset:} 
    A reduced version of the MNIST dataset, 
    containing only the digits 0 and 1, suitable 
    for binary classification tasks. This dataset 
    serves as a controlled environment to assess 
    the model's performance on simpler classification tasks.
\end{enumerate}

\subsection{Feature Reduction Methods}
We implemented and compared three distinct feature reduction approaches:
\subsubsection{Principal Component Analysis (PCA)}
PCA was employed as a baseline linear 
dimensionality reduction technique. 
We extracted the top-k principal components 
(typically 4-12) from the flattened image data, 
which were then used as inputs to both classical and 
quantum-enhanced classifiers.

\subsubsection{Autoencoder}
We implemented a neural network-based autoencoder with the following architecture:

\begin{itemize}
    \item An encoder network that compresses
    the input to a lower-dimensional latent space.

    \item A decoder network that reconstructs 
    the original input from the latent representation

    \item Batch normalization and dropout layers 
    to improve training stability and prevent overfitting.
\end{itemize}

The autoencoder was trained to minimize reconstruction 
loss using the Adam optimizer with learning rates 
between 0.001—0.01 and weight decay for regularization.

\subsubsection{Quantum-Guided Autoencoder}
We introduced a quantum-guided autoencoder 
that incorporates feedback from the quantum 
reservoir during training. This approach combines:

\begin{itemize}
    \item A standard autoencoder architecture
    \item A linear classifier that maps quantum embeddings to class labels.
    \item A classification loss computed on quantum embeddings
    \item A weighting parameter $\lambda$ to control the trade-off between 
    reconstruction and classification objectives~\ref{eq:composed_loss_function}.
    \item A differentiable surrogate model that approximates the quantum reservoir's behavior,
    enabling end-to-end backpropagation through the entire pipeline. The frequency 
    of quantum updates is controlled by a hyperparameter, allowing for
    flexible integration of quantum dynamics into the training process.
\end{itemize}

\subsubsection{Quantum Reservoir Computing Layer}
Our quantum reservoir computing implementation uses 
a Rydberg atom simulator with the following parameterizable
components:

\begin{itemize}
    \item \textbf{Atom Chain Length:} 
    The number of atoms in the chain, which can be adjusted 
    to control the reservoir's capacity and complexity.

    \item \textbf{Rabi Frequency:} 
    The global Rabi drive amplitude, which influences the 
    strength of interactions between atoms.

    \item \textbf{Detuning Parameters:} 
    Site-dependent detuning parameters that encode the input data.

    \item \textbf{Measurement Strategy:} 
    The specific observables measured to obtain quantum embeddings.

    \item \textbf{Encoding Scale:}
    The scale of the input data encoding, which can be adjusted
    to optimize the quantum reservoir's response to the input data.

    \item \textbf{Time steps:}
    The number of time steps for which the quantum reservoir evolves,
    allowing for the capture of temporal dynamics in the data.
\end{itemize}

\subsubsection{Classical Network Architectures}
After the traning process of the quantum-guided autoencoder,
the following quantum embeddings are passed to a linear classifier.
The classifier is a simple feedforward neural network with the following architecture:
\begin{itemize}
    \item A single hidden layer with a ReLU activation function.
    \item Cross-entropy loss for training, optimized using the Adam optimizer.
\end{itemize}

\subsection{Comparison Methods}
We benchmarked our model against some
classical methods, including:
\begin{itemize}
    \item \textbf{PCA + Linear Classifier:} 
    A linear classifier trained on the PCA-reduced features.

    \item \textbf{Autoencoder + Linear Classifier:} 
    A linear classifier trained on the latent representations from the autoencoder.

    \item \textbf{PCA + Neural Network:}
    A fully connected neural network trained on the PCA-reduced features.

    \item \textbf{Autoencoder + Neural Network:}
    A fully connected neural network trained on the latent representations from the autoencoder.
\end{itemize}

\subsection{Performance Metrics}
We evaluated the performance of our model using the following metrics:
\begin{itemize}
    \item \textbf{Classification Accuracy:} 
    The percentage of correctly classified instances in the test set.

    \item \textbf{F1 Score:} 
    The harmonic mean of precision and recall, providing a balanced measure of classification performance.

    \item \textbf{Confusion Matrix:} 
    A matrix summarizing the classification results, showing true positives, false positives, true negatives, and false negatives.

\end{itemize}

\subsection{Parameter Sweep Strategy}
To optimize the performance of our quantum-guided autoencoder, 
we conducted a systematic parameter sweep over the following hyperparameters:
\begin{itemize}
    \item \textbf{Guided Lambda Parameter (\( \lambda \))}: 
    This parameter controls the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function}. We tested values ranging from 0.1 to 0.9 in increments of 0.1.

    \item \textbf{Quantum Update Frequency:} 
    The frequency at which the quantum reservoir is updated during training. We experimented with update frequencies of 1, 5, and 10 epochs.

    \item \textbf{Quantum Parameters:} 
    Including the number of atoms in the chain, Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
    \item \textbf{Batch Size:}
    \item \textbf{Learning Rate:}
\end{itemize}
